---
title: "Homework V"
author: "Sarah Bertoni"
date: 20 February 2024
format:
  pdf: 
    fontsize: 9pt
    mainfont: Garamond
    sansfont: Garamond
editor: visual
output: 
  html_document:
    theme: cosmo
    latex_engine: pdflatex
    huxreg_options: "tabular={l|ccc}, tablewidth=0.1\\textwidth, tablefont=\\footnotesize"
    fig_caption: true
execute:
  echo: false
  warning: false
prefer-html: true
---

```{r}
library("modelsummary")
library("haven")
library("tidyverse")
library("kableExtra")
library("stargazer")
library("huxtable")
library("knitr")
library("spatstat")
library("gridExtra")
library(tinytable)
```

# Analyzing the French labor force

```{r}
lfs_11<-read_dta("lfshwk_fr11.dta")
lfs_12<-read_dta("lfshwk_fr12.dta")
```

## a) Definitions

The *13th International Conference of Labour Statisticians*[^1] adopted a resolution concerning concerning statistics of the economically active population, employment, unemployment and underemployment so as to set standards to all countries and allow for the usefulness of such standards in enhancing the international comparability of the statistics.

[^1]: https://www.ilo.org/resource/resolution-concerning-statistics-economically-active-population-employment-0

Employment and unemployment are defined as follow: - One is employed if he/she is above a specified age and during a certain reference period (either one day or one week) was either in paid employment or self-employment. - One is unemployed if he/she is above a specified age and during a certain reference period was without work, currently available for work and seeking work (i.e. had taken specific steps in a taken recent period to seek employment as above defined).

The labour force or "currently active population" consists of all persons who satisfy the requirements to be included either among the employed or the unemployed.

Conversely, the currently inactive population comprises all persons who were not employed or unemployed during the reference period (not economically active) because of either attendance at educational institutions, engagement in household duties, retirement or old age,or other reasons.

Following these definitions, we can determine the labour force participation rate and the unemployment rate:

-   **labor force participation rate** = $$\frac{\text{Economically active population}}{\text{Total population > same specified age}}$$
-   **unemployment rate** = $$\frac{\text{Unemployed population}}{\text{Economically active population}}$$

The unemployment rate is defined over the economically active population, while the labour force participation rate is defined over the total population over a certain age threshold that we will later establish.

## b) **Employment and Unemployment rates**

According to the ILO and INSEE we define:

-   **Employment rate** $$\frac{\text{Employed population}}{\text{Total population}}$$

Employment rate is usually computed for a subset of the population, the \*working age population\*\*. The working age population is differently identified depending on the definition, and can comprise people aged older than 15 years old (which in our dataset corresponds to the totality of the observations), people in the 15-65 years old range or people in the 15-75 range.

-   **Unemployment rate** $$\frac{\text{Unemployed population}}{\text{Economically active population}}$$

Unemployment rate is computed over the active population. Consequently, $\text{Employment rate + Unemployment rate} \neq 1$ since the denominators do not account for the same aggregate.

We go on to more precisely define employment to eventually calculate employment rate.

**The following are considered to be in employment within the meaning of the ILO:**

• persons aged 15-74 who, during the reference week:

◦ have "worked at least one hour, as an employee or self-employed"

◦ or who "work without pay with a member of their family";

◦ or who "had a sideline or a small paid job";

◦ or who nevertheless have a job from which they are absent under certain conditions of reason ;

• The following are considered to be in employment:

◦ without conditions: paid leave, sick leave, maternity or paternity leave, working time arrangements (flexible hours, compensatory rest, part-time, etc.), training authorized by the employer or as part of a work-study/apprenticeship contract, partial unemployment (including technical or bad weather);

◦ parental leave in the event of a planned absence of 3 months or less or receipt of compensatory income linked to the activity such as Prepare;

◦ off-season period in the context of a seasonal activity in the event of regular work in the context of this seasonal activity;

◦ other reason for absence in the event of a planned absence of 3 months or less.

**The following are not considered to be in employment:**

• parental leave lasting more than three months and without receipt of compensatory income linked to the activity such as Prepare;

• off-season period in the context of a seasonal activity, without regular work in the context of this seasonal activity;

• end of employment period;

• other reason in the event of a planned absence of more than 3 months.

• people aged 75 or over who "have a job, even for a few hours, as employees or self-employed"\
The variables are encoded as following according to the Insee dataset:

```{r}


variables <- data.frame(
  Variable = c("TRAREF", "PASTRA", "RABS:", "RABS (1)",
               "RABS (2)", "RABS (3)", "RABS (4)", "RABS (5)",
               "RABS (6)", "RABS (7)", "RABS (8)", "RABS (9)",
               "RABS (10)", "RABS (11)", "AIDFAM", "RATTRAP", "TEMP"),
  Definition = c("Has a job during the reference week.",
                 "Have a job from which they are absent under certain conditions of reason.",
                 "Reason for absence",
                 "Paid leave, including RTT days.",
                 "Sick leave, including child care, work accident.",
                 "Maternity or paternity leave, depending on gender.",
                 "Part-time work.",
                 "Parental leave.",
                 "Other unpaid leave.",
                 "Employer-paid training or training under an apprenticeship contract.",
                 "Partial unemployment, technical unemployment.",
                 "Suspension, end of employment period.",
                 "Strike.",
                 "Off-season period in a seasonal activity, or period preceding the start of employment.",
                 "Work without pay with a member of their family.",
                 "Had a sideline or a small paid job.",
                 "People aged 75 or over who have a job, even for a few hours, as employees or self-employed.")
)

kable(variables, format = "markdown", col.names = c("Variable", "Definition"), escape = FALSE, caption = "Variables and Definitions for Employment") %>%
  kable_classic_2(full_width = F) %>%
  column_spec(1, bold=T, border_right = T) %>%
  column_spec(2, width = "30em") %>% 
  row_spec(0, bold=T) 

```

```{r}
lfshwk_fr11 <- read_dta("lfshwk_fr11.dta")
#define variable for employement:
dataset <- lfshwk_fr11
dataset <- dataset %>% mutate(employed= case_when(traref==1~ 1,
                                           (pastra == 1 & rabs %in% c(1, 2, 3, 4, 7, 8, 10)) ~1,
                                           (pastra==1 & c(rabs== 5|rabs==6) &  rabsp <= 91)~1,
                        temp==1 ~1,
                        aidref == 1 ~ 1,
                         if_all(c(traref, pastra, rabs, rabsp, aidref, temp), is.na) ~ NA_integer_, .default = 0))
```

**The following are considered to be in employment within the meaning of the ILO:**

• are without a job during the reference week;

• are available for work in the two weeks following the reference week ;

• have taken an active job search in the last four weeks or have found a job starting within three months .

```{r}
variables_new <- data.frame(
  Variable = c("TRAREF", "DISPOC", "DEM","ULTJ"),
  Definition = c("Has a job during the reference week.",
                 "Available for work",
                 "Active job search", "Time period before the start of next job")
)

kable(variables_new, format = "markdown", col.names = c("Variable", "Definition"), escape = FALSE, caption = "Variables and Definitions for Unemployment") %>%
  kable_classic_2(full_width = F) %>%
  column_spec(1, bold=T, border_right = T) %>%
  column_spec(2, width = "30em") %>% 
  row_spec(0, bold=T) 



```

```{r}

dataset <- dataset %>% mutate(unemployed= case_when(employed==0 & dispoc == 1 & (dem==1|ultj< 91) ~1,
                         is.na(traref) ~ NA_integer_, .default = 0))
```

```{r}
lfshwk_fr12 <- read_dta("lfshwk_fr12.dta")
dataset12<- lfshwk_fr12

dataset12 <- dataset12 %>% mutate(employed= case_when(traref==1~ 1,
                                           (pastra == 1 & rabs %in% c(1, 2, 3, 4, 7, 8, 10)) ~1,
                                           (pastra==1 & c(rabs== 5|rabs==6) &  rabsp <= 91)~1,
                        temp==1 ~1,
                        aidref == 1 ~ 1,
                         if_all(c(traref, pastra, rabs, rabsp, aidref, temp), is.na) ~ NA_integer_, .default = 0))
dataset12 <- dataset12 %>% mutate(unemployed= case_when(employed==0 & dispoc == 1 & (dem==1|ultj< 91) ~1,
                         is.na(traref) ~ NA_integer_, .default = 0))
```

**Employment Rates**

Rates, shares and graphs are thereafter computed through the IINSEE definitions and dataset.

The employment rate is defined as the ratio between employed over the working age population, which we defined as the age range between 15 and 74 years old. We also compute the share with the more narrow definition of working age population as 15-65, and the rate on the total population.

```{r}
lm11_75<- lm(data=dataset[dataset$ag<75,], formula= employed ~1, weights = weight)
lm11_65<- lm(data=dataset[dataset$ag<65,], formula= employed ~1, weights = weight)
lm11_tot<- lm(data=dataset, formula= employed ~1, weights = weight)

lm12_65<- lm(data=dataset12[dataset12$ag<65,], formula= employed ~1, weights = weight)
lm12_75<- lm(data=dataset12[dataset12$ag<75,], formula= employed ~1, weights = weight)
lm12_tot<- lm(data=dataset12, formula= employed ~1, weights = weight)


names(lm11_65$coefficients)[1]<- "[age 15-65]"
names(lm11_75$coefficients)[1]<- "[age 15-75]"
names(lm11_tot$coefficients)[1]<- "[total population]"
names(lm12_65$coefficients)[1]<- "[age 15-65]"
names(lm12_75$coefficients)[1]<- "[age 15-75]"
names(lm12_tot$coefficients)[1]<- "[total population]"

rates_empl <- huxreg("2011 [age 15-65]" = lm11_65, 
                       "2011 [age 15-75]" = lm11_75,
                      "2011 [total population]" = lm11_tot,

                       "2012 [age 15-65]" = lm12_65,
                       "2012 [age 15-75]" = lm12_75,
                                           "2012 [total population]" = lm12_tot,

                  error_format = "({std.error})",
                  error_pos = "below",
                  statistics = c(N = "nobs"),
                  note = "{stars}",
                  align = "center") %>%
  set_caption("Employement Rates")


rates_empl_65 <- huxreg("2011" = lm11_65, "2012" = lm12_65,
       statistics = c(N = "nobs"))
rates_empl_75 <- huxreg("2011" = lm11_75, "2012" = lm12_75,
       statistics = c(N = "nobs"))
rates_empl_tot<- huxreg("2011" = lm11_tot, "2012" = lm12_tot,
       statistics = c(N = "nobs"))

table_4 <- rbind(rates_empl_65[1:3,], rates_empl_75[2:3,], rates_empl_tot[2:4,])
caption(table_4) <- c("Employement Rates")
set_latex_float(table_4, value = "H")

```

Employment rates computed on the working age population aged 15-65 are higher compared to the ones computed on the 15-75 aged population and the total population, in both reference years; however, they do not vary much - even do they are statistically different - since most individuals aged 64-99 are not active in the labor force.

Employment rates in 2011 are not statistically different compared to employment rates in 2012. However, it is still relevant to notice that the point estimate for employment rates is higher in 2012 compared to 2011 only when considering the 15-65 aged population, and lower otherwise; such difference could be explained by a fast aging population.

**Uneployment Rates**

The unemployment rate has to be computed over the economically active population, not on the total population, differently form the employment rate. We therefore define the economically active population as the sum of employed and unemployed.

```{r}
dataset <- dataset %>%  mutate(active=ifelse(employed==1|unemployed==1, 1, 0))
dataset12 <- dataset12 %>%  mutate(active=ifelse(employed==1|unemployed==1, 1, 0))
```

```{r}
lm11_75u<- lm(data=dataset[(dataset$ag<75 & dataset$active==1),], formula= unemployed ~1, weights = weight)
lm11_65u<- lm(data=dataset[(dataset$ag<65 & dataset$active==1),], formula= unemployed ~1, weights = weight)
lm11utot<- lm(data=dataset[( dataset$active==1),], formula= unemployed ~1, weights = weight)

lm12_65u<- lm(data=dataset12[(dataset12$ag<65& dataset12$active==1),], formula= unemployed ~1, weights = weight)
lm12_75u<- lm(data=dataset12[(dataset12$ag<75& dataset12$active==1),], formula= unemployed ~1, weights = weight)
lm12utot<- lm(data=dataset12[( dataset12$active==1),], formula= unemployed ~1, weights = weight)

names(lm11_65u$coefficients)[1]<- "[age 15-65]"
names(lm11_75u$coefficients)[1]<- "[age 15-75]"
names(lm11utot$coefficients)[1]<- "[total population]"
names(lm12_65u$coefficients)[1]<- "[age 15-65]"
names(lm12_75u$coefficients)[1]<- "[age 15-75]"
names(lm12utot$coefficients)[1]<- "[total population]"


rates_unempl_65 <- huxreg("2011" = lm11_65u, "2012" = lm12_65u,
       statistics = c(N = "nobs"))
rates_unempl_75 <- huxreg("2011" = lm11_75u, "2012" = lm12_75u,
       statistics = c(N = "nobs"))
rates_unempl_tot<- huxreg("2011" = lm11utot, "2012" = lm12utot,
       statistics = c(N = "nobs"))

table_5 <- rbind(rates_unempl_65[1:3,], rates_unempl_75[2:3,], rates_unempl_tot[2:4,])
caption(table_4) <- c("Unmployement Rates")
set_latex_float(table_5, value = "H")



```

Similarly to the case of employment, unmployment rates computed on the active population aged 15-65 are very close to the ones computed on the 15-75 aged population and the total population, in both reference years, and are not statistically different;

Unemployment rates in 2011 are statistically lower compared to unemployment rates in 2012.

## c) Own definition of unemployment

The official definition of unemployment has been often critised because it excludes from the labour force those who are not looking for work, ie. discouraged workers.

This relates to the broader problem of "unemployment halo" and the widespread criticism according to which ILO offical unemployment rates severely underestimate the "true" extent of unemployment. According to the insee definition, the unemployment halo consists of unemployed people who are either looking for work but are not available to work within two weeks, or who have not actively looked for work in the previous month but would like to work, whether or not they are available.

The halo therefore includes people who , while being part of the inactive population as per following of the official requirements defined by the ILO (neither employed nor unemployed), are close to the labour market.

To correct for such underestimation of the unemployment rate we computed ab adjusted rate: we include in the definition of unemployment people available for a job but not actively seeking, as they relate to people that may be discouraged in the reasearch but would accept a job if offered one.

We therefore re-calculate unemployment rates including individuals:

1\) seeking a job but not immediately available

2\) available for a job but not actively seeking

3\) not seeking and not immediately available

Our definition of unemployment is therefore simply the share of people over 15 years old who are involuntarily without work on the as such newly defined active population. We also elongate the period for the start of a future job to 300 days to account for people waiting for the next job. For semplicity, we focus on the working age population defined as in the 15-65 age range.

```{r}
dataset <- dataset %>% mutate(unemployed2= case_when(traref==2 & dispoc == 1 & (dem==1|ultj< 91) ~1, (employed==0 & dispoc == 0 & (dem==1|ultj< 91)) ~1, (employed==0 & dispoc == 0 & (dem==0|ultj< 91)) ~1, (employed==0 & dispoc == 0 & (dem==0|ultj< 300)) ~1,
                         is.na(traref) ~ NA_integer_, .default = 0))

dataset12 <- dataset12 %>% mutate(unemployed2= case_when(traref==2 & dispoc == 1 & (dem==1|ultj< 91) ~1, (employed==0 & dispoc == 0 & (dem==1|ultj< 300)) ~1, (employed==0 & dispoc == 0 & (dem==0|ultj< 91)) ~1, (employed==0 & dispoc == 0 & (dem==0|ultj< 300)) ~1,
                         is.na(traref) ~ NA_integer_, .default = 0))


dataset <- dataset %>%  mutate(activenew=ifelse(employed==1|unemployed2==1, 1, 0))
dataset12 <- dataset12 %>%  mutate(activenew=ifelse(employed==1|unemployed2==1, 1, 0))


lm11_65unew<- lm(data=dataset[(dataset$ag<65 & dataset$activenew==1),], formula= unemployed2 ~1, weights = weight)


lm12_65unew<- lm(data=dataset12[(dataset12$ag<65 & dataset12$activenew==1),], formula= unemployed2 ~1, weights = weight)

names(lm11_65unew$coefficients)[1]<- "Rate"
names(lm12_65unew$coefficients)[1]<- "Rate"


rates_unempl2 <- huxreg("2011" = lm11_65u, 
                      
                       "2012" = lm12_65u,
                     
                       
                       
                  error_format = "({std.error})",
                  error_pos = "below",
                  statistics = c(N = "nobs"),
                  note = "{stars}",
                  align = "center") 

rates_unempl_new <- huxreg("2011" = lm11_65unew, 
                       
                       
                       "2012" = lm12_65unew,
                      
                       
                       
                  error_format = "({std.error})",
                  error_pos = "below",
                  statistics = c(N = "nobs"),
                  note = "{stars}",
                  align = "center") %>% set_caption("Unemployment Rates Adjusted [age 15-65]")

table_5 <- cbind(rates_unempl2[2:3], rates_unempl_new[2:3])  

set_latex_float(rates_unempl_new, value = "H")

```

Our new definition, as expected, leads to statistically higher unemployment rates as it comprises an higher percentage of the job seekers. As we mention, the gap in unemployment rates may be caused by:

-   individuals not immediately available for a job (eg. because of personal or family reasons, completing education or training, sick or disabled).

-   individuals not actively seeking (eg. discouraged workers, persons unable to seek work, people waiting for interview results).

The adjusted rates more accurately reflect the labour market, as they account for the halo between employment and unemployment and between inactivity and unemployment.

## d) Standard errors

We assume that the LFS sample was drawn randomly: as such, we can compute the standard error of the unemployment rate knowing that the unemployment rate is a **proportion**, which follows a binomial distribution. Denoting $\hat{u}$ the estimator of the unemployment rate, its standard error is: $\hat{\sigma} = \sqrt\frac{\hat{u}(1-\hat{u})}{n-1}$

```{r}
x <- sqrt(0.092*(1-0.092)/nrow(dataset))
y <- sqrt(0.098*(1-0.098)/nrow(dataset12))
```

-   For 2011: we have $\hat{u}_{2011}$ = 0.092 Since $n= 423425$ observations: $$\hat{\sigma}_{2011} = \sqrt\frac{0.092\times0.908}{423424}$$ = `r round(x, 5)`.

-   For 2012: we have $\hat{u}_{2012}$ = 0.098 with $n= 422133$ $$\Longrightarrow \hat{\sigma}_{2012} = \sqrt\frac{0.098\times0.902}{422132}$$ = `r round(y, 5)`.

The resulting "95%%" confidence intervals $\text{CI} = \hat{u} \pm 1.96 \times \text{SE}(\hat{u})$ would account to: $$CI_{2011}=(0.0911,0.0929)$$ $$CI_{2012}=(0.0971,0.0989)$$

Assuming that the sample is random would lead to biased standard errors and artificially low standard errors. To account for differences in the representation of characteristics within the sample and allow to reflect the true confidence intervals of the estimates, sampling weights must be used.

## e) Evolution

To analyse the evolution of the composition of the working age population we compute unemployment and employment shares for each quarter, making up the active population; differently from the rates, unemployment shares are computed over the whole working age population, not only the active one. We define the inactive population as consisting of all those individuals who are neither employed nor unemployed and compute the relative share.

### 2011

```{r}
dataset <- dataset %>%  mutate(inactive=ifelse(employed==0 & unemployed==0, 1, 0))
dataset12 <- dataset12 %>%  mutate(inactive=ifelse(employed==0&unemployed==0, 1, 0))

lm11_1<- lm(data=dataset[(dataset$ag<65  & dataset$trim==1 ),], formula= unemployed ~1, weights = weight)


lm11_2<- lm(data=dataset[(dataset$ag<65 & dataset$trim==2 ),], formula= unemployed ~1, weights = weight)

lm11_3<- lm(data=dataset[(dataset$ag<65 & dataset$trim==3),], formula= unemployed ~1, weights = weight)

names(lm11_1$coefficients)[1]<- "Unemployment Share"
names(lm11_2$coefficients)[1]<- "Unemployment Share"
names(lm11_3$coefficients)[1]<- "Unemployment Share"



```

```{r}

lm11_65Q1<- lm(data=dataset[(dataset$ag<65 & dataset$trim==1 ),], formula= employed ~1, weights = weight)

lm11_65Q2<- lm(data=dataset[(dataset$ag<65 & dataset$trim==2 ),], formula= employed ~1, weights = weight)
lm11_65Q3<- lm(data=dataset[(dataset$ag<65 & dataset$trim==3 ),], formula= employed ~1, weights = weight)



names(lm11_65Q1$coefficients)[1]<- "Employment Share"

names(lm11_65Q2$coefficients)[1]<- "Employment Share"
names(lm11_65Q3$coefficients)[1]<- "Employment Share"


```

```{r}
lm11_1in<- lm(data=dataset[(dataset$ag<65  & dataset$trim==1 ),], formula= inactive ~1, weights = weight)


lm11_2in<- lm(data=dataset[(dataset$ag<65 & dataset$trim==2 ),], formula= inactive ~1, weights = weight)

lm11_3in<- lm(data=dataset[(dataset$ag<65 & dataset$trim==3 ),], formula= inactive ~1, weights = weight)


names(lm11_1in$coefficients)[1]<- "Inactive Share"
names(lm11_2in$coefficients)[1]<- "Inactive Share"
names(lm11_3in$coefficients)[1]<- "Inactive Share"



table_empl <- huxreg("Q1" = lm11_65Q1, "Q2" = lm11_65Q2, "Q3" = lm11_65Q3,
       statistics = c(N = "nobs"), note = NULL)
table_unempl <- huxreg("Q1" = lm11_1, "Q2" = lm11_2, "Q3" = lm11_3,
       statistics = c(N = "nobs"), note = NULL)
table_inac <- huxreg("Q1" = lm11_1in, "Q2" = lm11_2in, "Q3" = lm11_3in,
       statistics = c(N = "nobs"), note = NULL)


table_6 <- rbind(table_empl[1:3,], table_unempl[2:3,], table_inac[2:4,])
caption(table_6) <- c("2011 shares per Quarter [age 15-65]")
table_6
```

Overall, quarterly shares allow us to observe an increasing, while statistically not significant, trend in the employment share, compensated by a significant decreasing unemployment and a stable inactive population.

Between the first and second quarter unemployment significantly decreases from $6.8\%$ to $6.2\%$, while changes in the employment and inactive shares, which have to compensate the decrease in unemployment, are statistically no significant and therefore do not allow us to draw any conclusion on the direction of the shift; in particular, we cannot conclude whether the decrease in unemployment is due to people leaving the labour force or people become employed.

Between the second and third quarter there appear to be no significant change in the working age population composition; the point estimate for unemployment, however, increases, while inactive share decreases, which would mean that unemployment increases due to inactive people joining the labour force, most likely starting to search for a job.

```{r, fig.width=4, fig.height=3}

results <- data.frame(
  Quarter = rep(c("Q1", "Q2", "Q3"), times = 3),  # Trim values (Q1, Q2, Q3)
  Category = rep(c("employed", "unemployed", "inactive"), each = 3),
  Share = c(
    lm11_65Q1$coefficients[1],  # empl Q1
    lm11_65Q2$coefficients[1],  # empl Q2
    lm11_65Q3$coefficients[1],  # empl Q3
    lm11_1$coefficients[1],  # unempl Q1
    lm11_2$coefficients[1],  # unempl Q2
    lm11_3$coefficients[1],  # unempl Q3
    lm11_1in$coefficients[1],  # inac Q1
    lm11_2in$coefficients[1],  # inac Q2
    lm11_3in$coefficients[1]   # inac Q3
  )
)

#pivot long
results_long <- results %>%
  gather(key = "Variable", value = "Value", Share)

#  plot
ggplot(results_long, aes(x = factor(Quarter), y = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  labs(title = "Composition of working age population in 2011", 
       x = "Quarter", 
       y = "Share",
       fill = "Category") +
  scale_fill_manual(values = c("employed" = "beige", "unemployed" ="#918380", "inactive" = "#d2aea7")) +
  theme_minimal()+
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )+theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
     axis.title.x = element_text(size = 6),
     axis.title.y = element_text(size=6),
    legend.text = element_text(size = 6),
        legend.title = element_text(size = 7)) 
```

Overall, in 2011 we observe a quite stable quarterly composition of the working age population.

### 2012

```{r}
lm12_1 <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==1 ),], formula= unemployed ~1, weights = weight)
lm12_2 <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==2 ),], formula= unemployed ~1, weights = weight)

names(lm12_1$coefficients)[1] <- "Unemployment Share"
names(lm12_2$coefficients)[1] <- "Unemployment Share"

lm12_65Q1 <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==1 ),], formula= employed ~1, weights = weight)
lm12_65Q2 <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==2 ),], formula= employed ~1, weights = weight)

names(lm12_65Q1$coefficients)[1] <- "Employment Share"
names(lm12_65Q2$coefficients)[1] <- "Employment Share"

lm12_1in <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==1 ),], formula= inactive ~1, weights = weight)
lm12_2in <- lm(data=dataset12[(dataset12$ag<65 & dataset12$trim==2 ),], formula= inactive ~1, weights = weight)

names(lm12_1in$coefficients)[1] <- "Inactive Share"
names(lm12_2in$coefficients)[1] <- "Inactive Share"

#  combining the tables 
table_empl <- huxreg("Q1" = lm12_65Q1, "Q2" = lm12_65Q2,
       statistics = c(N = "nobs"), note = NULL)

table_unempl <- huxreg("Q1" = lm12_1, "Q2" = lm12_2,
       statistics = c(N = "nobs"), note = NULL)

table_inac <- huxreg("Q1" = lm12_1in, "Q2" = lm12_2in, 
       statistics = c(N = "nobs"), note = NULL)
table_6 <- rbind(table_empl[1:3,], table_unempl[2:3,], table_inac[2:4,])
caption(table_6) <- c("2012 Shares per Quarter [age 15-65]")
table_6
```

Similarly to 2011, employment share between the first and second quarter of 2012 increases (while being the decrease statistically not significant), while unemployment share appears to be statistically lower in the second quarter, with an overall stable inactive share, which would represent a shift from unemployment to employment, eg. job-seeking people finding an occupation.

```{r, fig.width=4, fig.height=3}

results12 <- data.frame(
  Quarter = rep(c("Q1", "Q2"), times = 3),  # Trim values (Q1, Q2)
  Category = rep(c("employed", "unemployed", "inactive"), each = 2),
  Share = c(
    lm12_65Q1$coefficients[1],  # empl Q1
    lm12_65Q2$coefficients[1],  # empl Q2
   
    lm12_1$coefficients[1],  # unempl Q1
    lm12_2$coefficients[1],  # unempl Q2
  
    lm12_1in$coefficients[1],  # inac Q1
    lm12_2in$coefficients[1])  # inac Q2
  
)

#pivot long
results12_long <- results12 %>%
  gather(key = "Variable", value = "Value", Share)

#  plot
ggplot(results12_long, aes(x = factor(Quarter), y = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  labs(title = "Composition of working age population in 2012", 
       x = "Quarter", 
       y = "Share",
       fill = "Category") +
  scale_fill_manual(values = c("employed" = "beige", "unemployed" ="#918380", "inactive" = "#d2aea7")) +
  theme_minimal()+
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )+theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
     axis.title.x = element_text(size = 6),
     axis.title.y = element_text(size=6),
    legend.text = element_text(size = 6),
        legend.title = element_text(size = 7)) 
```

## f) Unemployment rate and labor force participation by characteristics

We compute the unemployment rate and labor force participation by demographics, according to the official definition.

Men appear to experience a lower unemployment rate, even though such difference is not statistical significant; an higher unemployment rate could be associated with labour discrimination experienced by women, both for personal biases of the employeers and for leagl causes such as possible maternity leave.

On average, men are statistically more active in the labour force; such result can be explained by a higher share of women not pursuing a career in favour of taking care of the household, as well as extended or perment absence from the labour force due to maternity.

```{r}
# Gender
wmu <- lm(data=dataset12[(dataset12$ag<65 & dataset12$sexe==2 & dataset12$active==1),], formula= unemployed ~1, weights = weight)
menu <- lm(data=dataset12[(dataset12$ag<65 & dataset12$sexe==1 & dataset12$active==1),], formula= unemployed ~1, weights = weight)
wmlf <- lm(data=dataset12[(dataset12$ag<65 & dataset12$sexe==2 ),], formula= active ~1, weights = weight)
menlf <- lm(data=dataset12[(dataset12$ag<65 & dataset12$sexe==1 ),], formula= active ~1, weights = weight)

women <- huxreg("Unemployment" = wmu, "Labour Force" = wmlf,
       statistics = c(N = "nobs"), coefs = c("Women:" = "(Intercept)"), note = NULL)
men <- huxreg("Unemployment" = menu, "Labour Force" = menlf,
       statistics = c(N = "nobs"), coefs = c("Men:" = "(Intercept)"), note = NULL)


table_women <- rbind(women[1:3,], men[2:3,])
caption(table_women) <- c("Rates by Gender 2012")



results <- data.frame(
  Category = rep(c("Unemployed", "Labour Force"), each = 2),  # Categories for unemployed and labour force
  Gender = rep(c("Women", "Men"), times = 2),  # Gender
  Share = c(
    coef(wmu)[1],  # Women: Unemployment
    coef(menu)[1],  # Men: Unemployment
    coef(wmlf)[1],  # Women: Labour Force
    coef(menlf)[1]   # Men: Labour Force
  )
)

# Pivot the data to long format
results_long <- results %>%
  gather(key = "Variable", value = "Value", Share)

# Plot using ggplot2
plot1 <- ggplot(results_long, aes(x = Gender, y = (100*Value), fill = Category)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Unemployment and Labour Force Rates % by Gender", 
       x = "Gender", 
       y = "Rate (%)",
       fill = "Category") +
  scale_fill_manual(values = c("Unemployed" = "lightgoldenrod1", "Labour Force" = "burlywood")) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
    axis.title.x = element_text(size = 6),
    axis.title.y = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  )

set_latex_float(table_women, "H") 

```

Unemployment rates and partecipation rate vary widely across age groups, with every difference being statistically significant. The most active group is represented by the 25-50, as, intuitively, it corresponds to people who just entered the labour force after ending the education cycle - and does not include retirees.

Nonetheless, the middle group experience higher unemployment rate than the 50-64, which, conversely, is less active in the labour force, characteristic explained by retirement. The most striking observable result is represented by the high young unemployment rate, which paints a worrying picture of the labour market. Young people are less active in the labour force due to the high percentage of students.

```{r}
##age

ageuhigher <- lm(data=dataset12[((dataset12$ag>=50 & dataset12$ag<65 & dataset12$active==1)),], 
                 formula= unemployed ~1, weights = weight)

ageulower <- lm(data=dataset12[(dataset12$ag<24 & dataset12$active==1),], 
                formula= unemployed ~1, weights = weight)

ageumiddle <- lm(data=dataset12[((dataset12$ag>=24 & dataset12$ag<=50) & dataset12$active==1),], formula= unemployed ~1, weights = weight)

ageufhigher <- lm(data=dataset12[((dataset12$ag>=50 & dataset12$ag<65)),], 
                  formula= active ~1, weights = weight)

ageuflower <- lm(data=dataset12[(dataset12$ag<24),], 
                 formula= active ~1, weights = weight)

ageufmiddle <- lm(data=dataset12[((dataset12$ag>=24 & dataset12$ag<=50)),], 
                  formula= active ~1, weights = weight)

higher <- huxreg("Unemployment" = ageuhigher, "Labour Force" = ageufhigher,
       statistics = c(N = "nobs"), coefs = c("Age 50-64:" = "(Intercept)"), note = NULL)

lower <- huxreg("Unemployment" = ageulower, "Labour Force" = ageuflower,
       statistics = c(N = "nobs"), coefs = c("Age 15-24:" = "(Intercept)"), note = NULL)

middle <- huxreg("Unemployment" = ageumiddle, "Labour Force" = ageufmiddle,
       statistics = c(N = "nobs"), coefs = c("Age 25-50:" = "(Intercept)"), note = NULL)

# Combine the tables for output
table_age <- rbind(higher[1:3,], middle[2:3,],lower[2:4,])
caption(table_age) <- c("Rates by Age Group 2012")


# Create a data frame with coefficients for each age group
results_age <- data.frame(
  Category = rep(c("Unemployed", "Labour Force"), each = 3),  # Categories for unemployment and labour force
  AgeGroup = rep(c("Age 15-24", "Age 25-50", "Age 50-64"), times = 2),  # Age groups
  Share = c(
    coef(ageulower)[1],  # Age 15-24: Unemployment
    coef(ageumiddle)[1],  # Age 25-50: Unemployment
    coef(ageuhigher)[1],  # Age 50-64: Unemployment
    coef(ageuflower)[1],  # Age 15-24: Labour Force
    coef(ageufmiddle)[1],  # Age 25-50: Labour Force
    coef(ageufhigher)[1]   # Age 50-64: Labour Force
  )
)


results_age_long <- results_age %>%
  gather(key = "Variable", value = "Value", Share)


plot2 <- ggplot(results_age_long, aes(x = AgeGroup, y = (100 * Value), fill = Category)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Use "dodge" for side-by-side bars
  labs(title = "Unemployment and Labour Force Rates % by Age Group  2012", 
       x = "Age Group", 
       y = "Rate (%)",
       fill = "Category") +
  scale_fill_manual(values = c("Unemployed" = "lightgoldenrod1", "Labour Force" = "burlywood")) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
    axis.title.x = element_text(size = 6),
    axis.title.y = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  )

set_latex_float(table_age, "H") 

```

Unemployment and labour partecipation rates in 2012 highlights the positive returs to education in France, as people holding a degree higher than BAC experience significantly lower unemployment rates and higher partecipation in the labour force than those with an high school diploma or lower and those with a vocational diploma.

On the other hand, those in possession of a vocational diploma are less likely to be unemployed and on average more likely to be in the labour force than those without diploma or with an high school diploma.

The positive relationship between labour force participation and educational level may be explained by the high unemployment rates for the lower educated, which may lead many discouraged workers to leave the labour force.

```{r}

dataset12 <- dataset12 %>%
  mutate(
    Education = case_when(
      ddipl == 1 ~ "Higher than BAC",     # Higher than BAC
      ddipl == 3 ~ "High School or Lower", # BAC +2
      ddipl == 4 ~ "High School or Lower", # BAC
      ddipl == 5 ~ "Vocational Diploma",   # CAP/BEP (Vocational)
      ddipl == 6 ~ "High School or Lower", # Brevet des collèges
      ddipl == 7 ~ "High School or Lower", # Primary
      TRUE ~ "Unknown"                     # Handle unexpected values
    )
  )


age_diploma_high <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "Higher than BAC" & dataset12$active == 1),], 
                       formula= unemployed ~1, weights = weight)

age_diploma_voc <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "Vocational Diploma" & dataset12$active == 1),], 
                      formula= unemployed ~1, weights = weight)

age_diploma_low <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "High School or Lower" & dataset12$active == 1),], 
                      formula= unemployed ~1, weights = weight)

# Labour force models for the same categories
age_diploma_high_labour <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "Higher than BAC"),], 
                              formula= active ~1, weights = weight)

age_diploma_voc_labour <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "Vocational Diploma"),], 
                             formula= active ~1, weights = weight)

age_diploma_low_labour <- lm(data=dataset12[(dataset12$ag<65 & dataset12$Education == "High School or Lower"),], 
                             formula= active ~1, weights = weight)


results_diploma <- data.frame(
  Category = rep(c("Unemployed", "Labour Force"), each = 3),  # Categories for unemployment and labour force
  EducationLevel = rep(c("Higher than BAC", "Vocational Diploma", "High School or Lower"), times = 2),  
  
  Share = c(
    coef(age_diploma_high)[1],  # Higher than BAC: Unemployment
    coef(age_diploma_voc)[1],   # Vocational Diploma: Unemployment
    coef(age_diploma_low)[1],   # High School or Lower: Unemployment
    coef(age_diploma_high_labour)[1],  # Higher than BAC: Labour Force
    coef(age_diploma_voc_labour)[1],  # Vocational Diploma: Labour Force
    coef(age_diploma_low_labour)[1]   # High School or Lower: Labour Force
  )
)


results_diploma_long <- results_diploma %>%
  gather(key = "Variable", value = "Value", Share)


plot3 <- ggplot(results_diploma_long, aes(x = EducationLevel, y = (100 * Value), fill = Category)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  
  
  labs(title = "Unemployment and Labour Force Rates % by Education Level 2012", 
       x = "Education Level", 
       y = "Rates (%)",
       fill = "Category") +
  scale_fill_manual(values = c("Unemployed" = "lightgoldenrod1", "Labour Force" = "burlywood")) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
    axis.title.x = element_text(size = 6),
    axis.title.y = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  )


high_diploma <- huxreg("Unemployment" = age_diploma_high, "Labour Force" = age_diploma_high_labour,
                       statistics = c(N = "nobs"), coefs = c("Higher than BAC:" = "(Intercept)"), note = NULL)

vocational_diploma <- huxreg("Unemployment" = age_diploma_voc, "Labour Force" = age_diploma_voc_labour,
                             statistics = c(N = "nobs"), coefs = c("Vocational Diploma:" = "(Intercept)"), note = NULL)

low_diploma <- huxreg("Unemployment" = age_diploma_low, "Labour Force" = age_diploma_low_labour,
                     statistics = c(N = "nobs"), coefs = c("High School or Lower:" = "(Intercept)"), note = NULL)

# Combine the tables for output (only selecting relevant rows)
table_diploma <- rbind(
  high_diploma[1:3,],          # Select Unemployment and Labour Force for Higher than BAC
  vocational_diploma[2:3,],    # Select Unemployment and Labour Force for Vocational Diploma
  low_diploma[2:3,]           # Select Unemployment and Labour Force for High School or Lower
)

caption(table_diploma) <- c("Rates by Education Level 2012")


set_latex_float(table_diploma, "H")

```

To obtain picture of the difference in labour force participation and unemployment rates across demographic characteristics we plot the relative rates.

```{r, fig.width=4, fig.height=5}

results_diploma <- data.frame(
  Category = rep(c("Unemployed", "Labour Force"), each = 2),  
  
  EducationLevel = rep(c("Lower or no Diploma", "Higher Diploma"), times = 2), 
  
  Share = c(
    coef(age_diploma_low)[1],  # Low diploma: Unemployment
    coef(age_diploma_high)[1],  # High diploma: Unemployment
    coef(age_diploma_low_labour)[1],  # Low diploma: Labour Force
    coef(age_diploma_high_labour)[1]   # High diploma: Labour Force
  )
)

# Pivot the data to long format
results_diploma_long <- results_diploma %>%
  gather(key = "Variable", value = "Value", Share)

plot3 <- ggplot(results_diploma_long, aes(x = EducationLevel, y = (100 * Value), fill = Category)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Use "dodge" for side-by-side bars
  labs(title = "Unemployment and Labour Force Rates % by Education Level 2012", 
       x = "Education Level", 
       y = "Rate (%)",
       fill = "Category") +
  scale_fill_manual(values = c("Unemployed" = "lightgoldenrod1", "Labour Force" = "burlywood")) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  theme(
    plot.title = element_text(size = 8),     
    plot.subtitle = element_text(size = 7),  
    plot.caption = element_text(size = 7),
    axis.title.x = element_text(size = 6),
    axis.title.y = element_text(size = 6),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  )

grid.arrange(plot1, plot2, plot3, ncol=1)
```

```{r}

table8<- rbind(table_women[1:5,], table_age[2:5,], table_diploma[2:7,])
caption(table8) <- c("Rates by demographics")
table8
```

\newpage

## Complete Code

```{r ref.label=all_labels(), echo=T, eval=FALSE}


```
